---
title: "R Final Report: Coral Spawning"
author: "Daniel, Annie, & River"
date: "2024-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Setting Up 
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(leaflet)
library(ggplot2)
library(lubridate)
library(forcats)
library(tidyr)
library(emmeans)
library(RColorBrewer)
```

```{r}
raw.data <- read.csv("data/CSD_Tables.csv")
```
  
##### This report is generated based on the Indo-Pacific coral spawning dataset organized by Baird et al. (2021). Our basic understanding of the dataset and the formation of the research question is inspired by Lin et al. (2021).  

### Introduction
- Synchronized mass coral spawning is observed once a year worldwide and occurs days following a full moon.
- The lunar cycle (tide and moonlight) regulates environmental changes that can be detected by corals to coordinate mass spawning. 
- This dataset includes coral spawning observations in different eco-regions (Figure 1). 

<style>
.html-widget {
    margin: auto;
}
</style>
```{r, warning=F, echo=F, message=F, leaflet, fig.align="center", fig.cap="**Figure 1**: World map with all the observed coral spawning sites in 2016."}
split.data <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/") %>% #先把年月日拆開才能篩選2016
  filter(year == 2016) %>%
  mutate(
    date.x = as.Date(paste(year, month, day, sep = "-"), "%Y-%m-%d"),
    site.y = fct_reorder(Site, Latitude, .desc = FALSE)) %>% # 根據緯度重新排列地點
  group_by(date.x, site.y, Country) %>%  # 處理點點大小：把同天site同country的數據分在同一坨
  summarise(Count = n())# n()功能可以算算同一坨有多少個。接著命名為 Count

split.data2 <- split.data %>%
  left_join(
    raw.data %>%
      dplyr::select(Site, Latitude, Longitude) %>%
      distinct(), by = c("site.y" = "Site"))

i.want.set3 <- brewer.pal(length(unique(split.data$Country)), "Set3")

color_palette <- colorFactor(palette = "Set3", domain = levels(split.data2$Country))
leaflet(data = split.data2) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~Longitude,
    lat = ~Latitude,
    popup = ~site.y,
    fillOpacity = 1,
    radius = 6,
    weight = 0.5,
    color = "black",
    fillColor = ~color_palette(Country))
```
- The timing of synchronized mass coral spawning depends on whether coral gametes have matured yet. 
- Gametes maturity varies between coral species and potentially influenced by environmental factors like temperature. 
- We wonder whether there the coral spawning time is correlated with the latitude, where sites closer to the equator observe spawning earlier in the year. 
- Zooming in the scope, we would also like to compare the spawning time of different coral genus between Taiwan and Japan (Figure 2). 

<style>
.html-widget {
    margin: auto;
}
</style>
```{r, warning=F, echo=F, message=F, fig.align="center", fig.cap="**Figure 2**: Map of coral spawning sites observed in Taiwan and Japan."}
# select the countries needed (Taiwan & Japan)
spawning.jptw <- raw.data %>% 
  filter(Country %in% c('Japan', 'Taiwan')) %>%
  filter(Genus %in% c('Galaxea', 'Lobophyllia', 'Coelastrea', 'Dipsastraea', 'Favites', 'Goniastrea', 'Platygyra', 'Porites'))

# delete the unneeded columns
spawning.jptw[ , c('Subsite', 'O_n', 'Depth_m', 'N', 'No_start', 'Quality_start', 'No_end', 'Quality_end', 'Gamete_release', 'Situation', 'Timezone', 'Reference', 'Comments')] <- list(NULL)
spawning.jptw$color <- ifelse(spawning.jptw$Country == "Japan", "#F8766D", "#00BFC4")

# Mapping
leaflet() %>%
  addTiles() %>%
  setView(lng = 123, lat = 27, zoom = 4.5) %>% # Center the map on Taiwan
  addCircleMarkers(
    lng = spawning.jptw$Longitude,
    lat = spawning.jptw$Latitude,
    popup = spawning.jptw$Site,
    radius = 5,
    fillColor = spawning.jptw$color,
    fillOpacity = 1,
    stroke = F)
```

### Methods
#### Spawning time and geolocation
We downloaded the Indo-Pacific coral spawning database (Baird et al. 2021) from nature.com/sdata/. Using the `max` and `summary` functions, it is determined that the year 2016 has the most data entry. Using the `ggplot2` package, we plot the spawning data (date and location) from 2016. 

```{r, warning=F}
split <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/")
summary(as.factor(split$year))
max(summary(as.factor(split$year))) # deternine most data year
```

##### Spawning data from Taiwan and Japan
Using the `filter` function, we selected the genus that have spawning data in both Taiwan and Japan. We then made a boxplot to compare their *date of spawning relative to nearest full moon (DoSRtNFM).   

Normality, Wilcoxon 
```{r}
# select the countries needed (Taiwan & Japan)
spawning.jptw <- raw.data %>% 
  filter(Country %in% c('Japan', 'Taiwan')) %>%
  filter(Genus %in% c('Galaxea', 'Lobophyllia', 'Coelastrea', 'Dipsastraea', 'Favites', 'Goniastrea', 'Platygyra', 'Porites'))

# delete the unneeded columns
spawning.jptw[ , c('Subsite', 'O_n', 'Depth_m', 'N', 'No_start', 'Quality_start', 'No_end', 'Quality_end', 'Gamete_release', 'Situation', 'Timezone', 'Reference', 'Comments')] <- list(NULL)
spawning.jptw$color <- ifelse(spawning.jptw$Country == "Japan", "#F8766D", "#00BFC4")
```

### Results
#### Spawning time and geolocation
result here
```{r, warning=F, echo=F, message=F, fig.align='center', fig.cap="**Figure 3**:_________"}
split.data <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/") %>% #先把年月日拆開才能篩選2016
  filter(year == 2016) %>%
  mutate(
    date.x = as.Date(paste(year, month, day, sep = "-"), "%Y-%m-%d"),
    site.y = fct_reorder(Site, Latitude)) %>% # 根據緯度重新排列地點
  group_by(date.x, site.y, Country) %>%  # 處理點點大小：把同天site同country的數據分在同一坨
  summarise(Count = n())# n()功能可以算算同一坨有多少個。接著命名為 Count

split.data2 <- split.data %>%
  left_join(
    raw.data %>%
      dplyr::select(Site, Latitude, Longitude) %>%
      distinct(), by = c("site.y" = "Site"))

i.want.set3 <- brewer.pal(length(unique(split.data$Country)), "Set3")
#set1_colors：指定點點顏色為Set3！
#unique()：獲取該欄位中所有不重複的值，也就是所有不同的國家。
#length()：計算不重複值的數量，就是總共有幾個國家
#brewer.pal(n, "Set3")：從 RColorBrewer 的 Set3 調色板中提取 n 種顏色

ggplot(split.data2, aes(x =date.x, y = site.y, size = Count, color = Country)) +
  geom_point(alpha = 0.4) +  
  scale_size_continuous(name = "Count", range = c(2, 8)) +  # 調整點大小
  scale_color_manual(values = i.want.set3) +  
  scale_x_date(date_labels = "%m/%d", date_breaks = "1 month") +
  labs(
    title = "Coral Spawning Observed in 2016",
    x = "Month",
    y = "Site",
    color = "Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        scale_fill_discrete(name = "Country")) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))

split.data2$Country <- factor(split.data2$Country)
```
*****
Notes   
World spawning figure:  
-	I think the latitude order is messed up? Bitoujiao shouldn’t be at the bottom?   
-	Change title to “Coral spawning observed in 2016”  

##### Spawning data from Taiwan and Japan
result here 
```{r, warning=F, echo=F, message=F, fig.align='center', fig.cap="**Figure 4**:_________"}
genus.tj.day <- ggplot(data = spawning.jptw, aes(x = Genus, y = DoSRtNFM, color = Country)) + 
  geom_boxplot() + 
  ggtitle("Spawning Time Comparison", subtitle = "different genus in Japan and Taiwan") + 
  ylab("DoSRtNFM*") + 
  theme(
    plot.title = element_text(size=16),
    plot.subtitle = element_text(size=12))
genus.tj.day
```
##### Stats
```{r, warning=F}
#check normality----
par(mfrow = c(1, 2))
hist(spawning.jptw$DoSRtNFM)
qqnorm(spawning.jptw$DoSRtNFM, main = 'Normal')
qqline(spawning.jptw$DoSRtNFM)
```
```{r}
#check data number----
J.genus <- spawning.jptw %>% 
          filter(Country == "Japan") %>% 
          dplyr::select(Genus)
summary(as.factor(J.genus$Genus))
T.genus <- spawning.jptw %>% 
  filter(Country == "Taiwan") %>% 
  dplyr::select(Genus)
summary(as.factor(T.genus$Genus))
```
sample size of *Lobophyllia* is too small, so there's no comparison below

***

```{r, warning=F}
#wilcoxon----
Coelastrea <- spawning.jptw %>%
  filter(Genus %in% "Coelastrea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Coelastrea)

Dipsastraea <- spawning.jptw %>%
  filter(Genus %in% "Dipsastraea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Dipsastraea)

Favites <- spawning.jptw %>%
  filter(Genus %in% "Favites",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Favites)

Galaxea <- spawning.jptw %>%
  filter(Genus %in% "Galaxea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Galaxea)

Goniastrea <- spawning.jptw %>%
  filter(Genus %in% "Goniastrea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Goniastrea)

Platygyra <- spawning.jptw %>%
  filter(Genus %in% "Platygyra",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Platygyra)

Porites <- spawning.jptw %>%
  filter(Genus %in% "Porites",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Porites)
```


### Discussion
Interpret the results here


### References 
Baird AH, Guest JR, Edwards AJ, Bauman AG, Bouwmeester J, Mera H, Abrego D, Alvarez-Noriega M, Babcock RC, Barbosa MB, et al. 2021. An Indo-Pacific coral spawning database. Sci Data. 8(1):35. doi:10.1038/s41597-020-00793-8.  

Lin C-H, Takahashi S, Mulla AJ, Nozawa Y. 2021. Moonrise timing is key for synchronized spawning in coral Dipsastraea speciosa. Proc Natl Acad Sci USA. 118(34):e2101985118. doi:10.1073/pnas.2101985118.









```{r}
split.data <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/") %>% #先把年月日拆開才能篩選2016
  filter(year == 2016) %>%
  mutate(
    date.x = as.Date(paste(year, month, day, sep = "-"), "%Y-%m-%d"),
    site.y = fct_reorder(Site, Latitude, .desc = FALSE)) %>% # 根據緯度重新排列地點
  group_by(date.x, site.y, Country) %>%  # 處理點點大小：把同天site同country的數據分在同一坨
  summarise(Count = n())# n()功能可以算算同一坨有多少個。接著命名為 Count

split.data2 <- split.data %>%
  left_join(
    raw.data %>%
      dplyr::select(Site, Latitude, Longitude) %>%
      distinct(), by = c("site.y" = "Site"))

i.want.set3 <- brewer.pal(length(unique(split.data$Country)), "Set3")
#set1_colors：指定點點顏色為Set3！
#unique()：獲取該欄位中所有不重複的值，也就是所有不同的國家。
#length()：計算不重複值的數量，就是總共有幾個國家
#brewer.pal(n, "Set3")：從 RColorBrewer 的 Set3 調色板中提取 n 種顏色

ggplot(split.data, aes(x =date.x, y = site.y, size = Count, color = Country)) +
  geom_point(alpha = 0.4) +  
  scale_size_continuous(name = "Count", range = c(2, 8)) +  # 調整點大小
  scale_color_manual(values = i.want.set3) +  
  scale_x_date(date_labels = "%m/%d", date_breaks = "1 month") +
  labs(
    title = "Sampling Distribution by Date and Site in 2016",
    x = "Date (2016)",
    y = "Site (ordered by latitude)",
    color = "Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        scale_fill_discrete(name = "Country")) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))

split.data2$Country <- factor(split.data2$Country)
```

```{r}
# Figure 1: World map with all the observed coral spawning sites.
color_palette <- colorFactor(palette = "Set3", domain = levels(split.data2$Country))
leaflet(data = split.data2) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~Longitude,
    lat = ~Latitude,
    popup = ~site.y,
    fillOpacity = 1,
    radius = 6,
    weight = 0.5,
    color = "black",
    fillColor = ~color_palette(Country))
```

```{r}
# select the countries needed (Taiwan & Japan)
spawning.jptw <- raw.data %>% 
  filter(Country %in% c('Japan', 'Taiwan')) %>%
  filter(Genus %in% c('Galaxea', 'Lobophyllia', 'Coelastrea', 'Dipsastraea', 'Favites', 'Goniastrea', 'Platygyra', 'Porites'))

# delete the unneeded columns
spawning.jptw[ , c('Subsite', 'O_n', 'Depth_m', 'N', 'No_start', 'Quality_start', 'No_end', 'Quality_end', 'Gamete_release', 'Situation', 'Timezone', 'Reference', 'Comments')] <- list(NULL)
spawning.jptw$color <- ifelse(spawning.jptw$Country == "Japan", "#F8766D", "#00BFC4")
```

```{r}
genus.tj.day <- ggplot(data = spawning.jptw, aes(x = Genus, y = DoSRtNFM, color = Country)) + 
  geom_boxplot() + 
  ggtitle("Spawning Time Comparison", subtitle = "different genus in Japan and Taiwan") + 
  ylab("DoSRtNFM*") + 
  theme(
    plot.title = element_text(size=20),
    plot.subtitle = element_text(size=12))
genus.tj.day
```
  
*Date of Spawning Relative to Nearest Full Moon
```{r}
#check normality----
par(mfrow = c(1, 2))
hist(spawning.jptw$DoSRtNFM)
qqnorm(spawning.jptw$DoSRtNFM, main = 'Normal')
qqline(spawning.jptw$DoSRtNFM)
```
```{r}
#check data number----
J.genus <- spawning.jptw %>% 
          filter(Country == "Japan") %>% 
          dplyr::select(Genus)
summary(as.factor(J.genus$Genus))
T.genus <- spawning.jptw %>% 
  filter(Country == "Taiwan") %>% 
  dplyr::select(Genus)
summary(as.factor(T.genus$Genus))
```
sample size of *Lobophyllia* is too small, so there's no comparison below

***

```{r}
#wilcoxon----
Coelastrea <- spawning.jptw %>%
  filter(Genus %in% "Coelastrea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Coelastrea)

Dipsastraea <- spawning.jptw %>%
  filter(Genus %in% "Dipsastraea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Dipsastraea)

Favites <- spawning.jptw %>%
  filter(Genus %in% "Favites",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Favites)

Galaxea <- spawning.jptw %>%
  filter(Genus %in% "Galaxea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Galaxea)

Goniastrea <- spawning.jptw %>%
  filter(Genus %in% "Goniastrea",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Goniastrea)

Platygyra <- spawning.jptw %>%
  filter(Genus %in% "Platygyra",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Platygyra)

Porites <- spawning.jptw %>%
  filter(Genus %in% "Porites",
         Country %in% c("Taiwan", "Japan"))
wilcox.test(DoSRtNFM ~ Country, data = Porites)
```

```{r}
split <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/")
summary(as.factor(split$year))
max(summary(as.factor(split$year)))
```

```{r}
split.data <- raw.data %>%
  separate(Date, into = c("day", "month", "year"), sep = "/") %>% #先把年月日拆開才能篩選2016
  filter(year == 2016) %>%
  mutate(
    date.x = as.Date(paste(year, month, day, sep = "-"), "%Y-%m-%d"),
    site.y = fct_reorder(Site, Latitude, .desc = FALSE)) %>% # 根據緯度重新排列地點
  group_by(date.x, site.y, Country) %>%  # 處理點點大小：把同天site同country的數據分在同一坨
  summarise(Count = n())# n()功能可以算算同一坨有多少個。接著命名為 Count

split.data2 <- split.data %>%
  left_join(
    raw.data %>%
      dplyr::select(Site, Latitude, Longitude) %>%
      distinct(), by = c("site.y" = "Site"))

i.want.set3 <- brewer.pal(length(unique(split.data$Country)), "Set3")
#set1_colors：指定點點顏色為Set3！
#unique()：獲取該欄位中所有不重複的值，也就是所有不同的國家。
#length()：計算不重複值的數量，就是總共有幾個國家
#brewer.pal(n, "Set3")：從 RColorBrewer 的 Set3 調色板中提取 n 種顏色

ggplot(split.data2, aes(x =date.x, y = site.y, size = Count, color = Country)) +
  geom_point(alpha = 0.4) +  
  scale_size_continuous(name = "Count", range = c(2, 8)) +  # 調整點大小
  scale_color_manual(values = i.want.set3) +  
  scale_x_date(date_labels = "%m/%d", date_breaks = "1 month") +
  labs(
    title = "Sampling Distribution by Date and Site in 2016",
    x = "Date (2016)",
    y = "Site (ordered by latitude)",
    color = "Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        scale_fill_discrete(name = "Country")) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))

split.data2$Country <- factor(split.data2$Country)
```

```{r}
# Figure 1: World map with all the observed coral spawning sites.
color_palette <- colorFactor(palette = "Set3", domain = levels(split.data2$Country))
leaflet(data = split.data2) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~Longitude,
    lat = ~Latitude,
    popup = ~site.y,
    fillOpacity = 1,
    radius = 6,
    weight = 0.5,
    color = "black",
    fillColor = ~color_palette(Country))
```



